---
title: "IB Report Card Analysis"
author: "Dave Rosborough"
date: "Nov 24, 2015"
output: html_document
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(tools)
```

The purpose of this analysis is to track students' anticipated grades and final results through their two year programme.

## Loading and Preparing the Data

The data is stored in two folders: `managebac` contains CSV exports for each reporting period.  `ibis` contains the CSV export from IBIS of final results.

First, we load all of the CSV files from the managebac directory, and combine their results into a single data frame.

```{r}

datadir <- "managebac"
datafiles <- list.files(path=datadir, pattern="*.csv", full.names=TRUE)
term_names <- file_path_sans_ext(basename(datafiles))
term_info <- do.call(rbind, strsplit(term_names, "-"))


mb_data <- lapply(datafiles, read.csv, header=FALSE, skip=1, na.strings = c("NA", "N/A", "IP"), col.names=c("class_id", "student_id", "last_name", "first_name", "average", "final_grade", "current_grade", "work_habits"))

mb_data <- mapply(cbind, mb_data, term = paste(term_info[,1], term_info[,2], sep="-"), SIMPLIFY=FALSE)

mb_data <- do.call(rbind, mb_data)

mb_data <- mb_data[!((is.na(mb_data$current_grade)) & (is.na(mb_data$work_habits))),]

mb_data <- subset(mb_data, select = -c(average, final_grade))

# remove section id to combine students who switched sections between years
mb_data$class_id <- as.factor(gsub(" 0[12]$", "", mb_data$class_id))

rm(term_info, datadir, datafiles, term_names)

mb_data.reshaped <- reshape(mb_data, direction="wide", timevar="term", idvar=c("class_id", "student_id", "last_name", "first_name"))

write.csv(mb_data.reshaped, "out.csv")
```

